---
title: "Learning R with Group - Week 4"
author: "Binod"
date: "4/26/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```

## Animal Survey Data set

```{r}
download.file(url="https://ndownloader.figshare.com/files/2292169",
              destfile = "data/survey.csv")
```


Column |	Description
--- | ---
record_id	 | Unique id for the observation
month	| month of observation
day	| day of observation
year	| year of observation
plot_id	| ID of a particular plot
species_id	| 2-letter code
sex	| sex of animal (“M”, “F”)
hindfoot_length	| length of the hindfoot in mm
weight	| weight of the animal in grams
genus	| genus of animal
species	| species of animal
taxon	| e.g. Rodent, Reptile, Bird, Rabbit
plot_type	| type of plot

```{r}
library(readr)

surveys <- read_csv("data/survey.csv")
```

## Structure

```{r}
str(surveys)
```

## Head / Tail

```{r}
head(surveys)
```

## Summary

```{r}
summary(surveys)
```

## Factors

- Factors represent categorical data. 

```{r}
sex <- factor(c("male", "female", "female", "male"))
```

### Levels

```{r}
levels(sex)  # name of levels

nlevels(sex) #number of levels
```

```{r}
sex # current order
```
### Ordering

```{r}
sex <- factor(sex, levels = c("male", "female"))
sex # after re-ordering
```

### Converting factors

```{r}
as.character(sex)

as.factor(sex)
```
### Year factorize

```{r}
year_fct <- factor(c(1990, 1983, 1977, 1998, 1990))
as.numeric(year_fct)               # Wrong! And there is no warning...
as.numeric(as.character(year_fct)) # Works...
as.numeric(levels(year_fct))[year_fct]    # The recommended way.
```
### Renaming factors

```{r}
plot(surveys$sex)
```

### Head

```{r}
sex <- surveys$sex
head(sex)
```

## Levels

```{r}
levels(sex)

```

```{r}
levels(sex)[1] <- "NA"
levels(sex)
```
## Date formatting

```{r}
str(surveys)
```

```{r}
library(lubridate)
```

### Create a new date

```{r}
new_date <- ymd("2015-01-01")
str(new_date)


# mdy()
# dmy()
```
### Merge Dates

```{r}
# paste() - combine two entity or columns

surveys$date <- ymd(paste(surveys$year, surveys$month, surveys$day, sep = "-"))

surveys$date
```

```{r}
summary(surveys$date)
```
## Missing dates

```{r}
missing_dates <- surveys[is.na(surveys$date), c("year", "month", "day")]

head(missing_dates)
```

# Data Manipulation in R


```{r}
surveys <- read_csv("data/survey.csv")
```

```{r}
str(surveys)
```

```{r}
View(surveys)
```

## Dplyr functions

- select(): subset columns
- filter(): subset rows on conditions
- mutate(): create new columns by using information from other columns
- group_by() and summarize(): create summary statistics on grouped data
- arrange(): sort results
- count(): count discrete values

## Selecting columns and filtering rows

```{r}
library(dplyr)
names(surveys)

select(surveys, plot_id, species_id, weight, sex)
```

### Deselect columns

```{r}
select(surveys, -record_id, -species_id) ## using - for deselect  
```

### Filter rows

```{r}
filter(surveys, year == 1978, sex == "M")
```

## Using %>% pipes %>% (ctrl + shift + m)

```{r}
surveys %>%
  filter(year == 1978) %>%
  select(species_id, sex, weight, year)
```
## Save the result

```{r}
surveys_sml <- surveys %>%
  filter(weight < 5) %>%
  select(species_id, sex, weight)

```

```{r}
surveys_sml
```

## Mutate - creating new variable

```{r}
surveys %>%
  mutate(weight_kg = weight / 1000)
```

```{r}
surveys %>%
  mutate(weight_kg = weight / 1000,
         weight_lb = weight_kg * 2.2)
```


```{r}
surveys %>%
  mutate(weight_kg = weight / 1000) %>%
  head()
```


### Remove missing value
```{r}
surveys %>%
  filter(!is.na(weight)) %>%
  mutate(weight_kg = weight / 1000) %>%
  head()
```
## Grouping and Summarize data

```{r}
surveys %>%
  group_by(sex) %>%
  summarize(mean_weight = mean(weight, na.rm = TRUE))
```

```{r}
surveys %>%
  group_by(sex, species_id, plot_id) %>%
  summarize(mean_weight = mean(weight, na.rm = TRUE))
```

## Arrange

```{r}
surveys %>%
  filter(!is.na(weight)) %>%
  group_by(sex, species_id) %>%
  summarize(mean_weight = mean(weight),
            min_weight = min(weight)) %>%
  arrange(desc(mean_weight))
```

## Counting

```{r}
surveys %>%
    count(sex) 
```
## Similar to above

```{r}
surveys %>%
    group_by(sex) %>%
    summarise(count = n())
```
## Sort / Arrange in count

```{r}
surveys %>%
    count(weight, sort = F) 
```

```{r}
surveys %>%
  count(sex, species) 
```
## Arrange

```{r}
surveys %>%
  count(sex, species) %>%
  arrange(species, desc(n))
```

```{r}
surveys_gw <- surveys %>%
  filter(!is.na(weight)) %>%
  group_by(genus, plot_id) %>%
  summarize(mean_weight = mean(weight))

str(surveys_gw)
```

## Exporting data

```{r}
surveys_complete <- surveys %>%
  filter(!is.na(weight),           # remove missing weight
         !is.na(hindfoot_length),  # remove missing hindfoot_length
         !is.na(sex))                # remove missing sex
```

## Saving Data

```{r}
surveys_complete

write_csv(surveys_complete, "./data/surveys_complete.csv")
```

