---
title: "Learning R with Group - Week 6"
author: "Binod"
date: "4/26/2020"
output: html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```

## Animal Survey Data set

```{r}
library(tidyverse)

# surveys_complete <- read_csv(file.choose())

surveys_complete <- read_csv("data/surveys_complete.csv")
```


Column |	Description
--- | ---
record_id	 | Unique id for the observation
month	| month of observation
day	| day of observation
year	| year of observation
plot_id	| ID of a particular plot
species_id	| 2-letter code
sex	| sex of animal (“M”, “F”)
hindfoot_length	| length of the hindfoot in mm
weight	| weight of the animal in grams
genus	| genus of animal
species	| species of animal
taxon	| e.g. Rodent, Reptile, Bird, Rabbit
plot_type	| type of plot


### Structure

```{r}
str(surveys_complete)
```

## ggplot2 - grammar of graphics
 
> ggplot(data = <data>, mapping = aes(<axis>)) +  <geom_function>()

## Load Data

```{r}
ggplot(data = surveys_complete)
```

## Add aesthetics

```{r}
ggplot(data = surveys_complete, aes(x = weight, y = hindfoot_length))
```

## Add geom 

- It's graphical representations of points

* `geom_point()` for scatter plots, dot plots, etc.
* `geom_boxplot()` boxplots
* `geom_line()` for trend lines, time series, etc.  

```{r}
ggplot(data = surveys_complete, 
       mapping = aes(x = weight, y = hindfoot_length)) +
  geom_point(size = 0.2) +
  labs(title = "Tile", subtitle = "Subtile", caption = "Source")
```

# Assign plot to a variable

```{r}
surveys_plot <- ggplot(data = surveys_complete, 
                       mapping = aes(x = weight, y = hindfoot_length))

# Draw the plot
surveys_plot + 
    geom_point() 
    
```
 ## Add transparency (alpha) to avoid overplotting

```{r}
surveys_plot + 
    geom_point(alpha = 0.1, size = 0.5)
```
## Adding color

```{r}
surveys_plot + 
    geom_point(alpha = 0.1, color = "red", size = 0.4)
```

## Colorize by Category


```{r}
surveys_plot + 
    geom_point(alpha = 0.5, aes(color = species_id), size = 0.5)
```
## Jitter

```{r}
surveys_plot +
    geom_jitter(alpha = 0.3, aes(color = species_id))
```
## Boxplot

We can use boxplots to visualize the distribution of weight within each species:

```{r}
ggplot(data = surveys_complete, 
       mapping = aes(x = species_id, y = weight)) +
    geom_boxplot()

# survey_plot <- ggplot(data = surveys_complete, mapping = aes(x = species_id, y = weight))

```
## Boxplot and Jitter

```{r}
surveys_plot +
    geom_boxplot(alpha = 0) +
    geom_jitter(alpha = 0.3, size = 0.5, color = "red")
```

## Plotting time series data

* Let’s calculate number of counts per year for each genus. 
* First we need to group the data and count records within each group:


```{r}
yearly_counts <- surveys_complete %>%
  count(year, genus)

# yearly_counts
```
## Yearlt line chart

```{r}
ggplot(data = yearly_counts, 
       mapping = aes(x = year, y = n)) +
     geom_line()
```


```{r}
ggplot(data = yearly_counts, 
       mapping = aes(x = year, y = n, group = genus)) +
    geom_line()
```

```{r}
ggplot(data = yearly_counts, 
       mapping = aes(x = year, y = n, color = genus)) +
    geom_line()
```
## Faceting

It allows the user to split one plot into multiple plots based on a factor included in the dataset.

There are two types of facet functions:

* facet_wrap() - arranges a 1-D sequence of panels 
* facet_grid() - form a matrix of rows and columns of panels.


### Facet wrap

```{r}
ggplot(data = yearly_counts, mapping = aes(x = year, y = n)) +
    geom_line() +
    facet_wrap(facets = vars(genus))
```

## Yearly sex counts

```{r}
yearly_sex_counts <- surveys_complete %>%
  count(year, genus, sex)
```


### Faceted plot by splitting further by sex using color (within each panel

```{r}
facet_yearly_sex_counts <- ggplot(data = yearly_sex_counts, 
       mapping = aes(x = year, y = n, color = sex))

facet_yearly_sex_counts +
  geom_line() +
  facet_wrap(facets =  vars(genus))
```

## Facet Grid

Using facet_grid() to control how panels are organized (rows and columns)


```{r}
facet_yearly_sex_counts +
  geom_line() +
  facet_grid(rows = vars(sex), cols =  vars(genus))
```
You can also organise the panels only by rows (or only by columns):

```{r}
# One column, facet by rows
ggplot(data = yearly_sex_counts, 
       mapping = aes(x = year, y = n, color = sex)) +
  geom_line() +
  facet_grid(rows = vars(genus))
```


```{r}
# One row, facet by column
facet_yearly_sex_counts +
  geom_line() +
  facet_grid(cols = vars(genus))
```

## ggplot2 themes

```{r}
 facet_yearly_sex_counts +
     geom_line() +
     facet_wrap(vars(genus)) +
     theme_bw()  # themes
```

## Do it yourself

Create a plot that depicts how the average weight of each species changes through the years.

```{r}
yearly_weight <- surveys_complete %>%
                group_by(year, species_id) %>%
                 summarize(avg_weight = mean(weight))
ggplot(data = yearly_weight, mapping = aes(x=year, y=avg_weight)) +
   geom_line() +
   facet_wrap(vars(species_id)) +
   theme_bw()
```

## Customize  

```{r}
p_yearly_sex_count <- facet_yearly_sex_counts +
    geom_line() +
    facet_wrap(vars(genus)) +
    labs(title = "Observed genera through time",
         x = "Year of observation",
         y = "Number of individuals") +
    theme_bw()

p_yearly_sex_count 
```

## Readability - Increase font size

```{r}
p_yearly_sex_count +
    theme(text = element_text(size = 16))
```
## Rotate the labels of x-axis

```{r}
# install.packages("extrafont") # for windows users

p_yearly_sex_count +
    theme(axis.text.x = element_text(colour = "grey20", 
                                     size = 12, angle = 90, 
                                     hjust = 0.5, vjust = 0.5),
         axis.text.y = element_text(colour = "grey20", size = 12),
         text = element_text(size = 16))
```

## Define custom theme

```{r}
# define custom theme
grey_theme <- theme(axis.text.x = element_text(colour = "grey20", size = 12, angle = 90, 
                                               hjust = 0.5, vjust = 0.5),
                          axis.text.y = element_text(colour = "grey20", size = 12),
                          text = element_text(size = 16))

# create a boxplot with the new theme
ggplot(surveys_complete, aes(x = species_id, y = hindfoot_length)) +
    geom_boxplot() +
    grey_theme
```


### Combine plots

```{r}
# install.packages("gridExtra")

library(gridExtra)

spp_weight_boxplot <- ggplot(data = surveys_complete, 
                             mapping = aes(x = genus, y = weight)) +
  geom_boxplot() +
  scale_y_log10() +  ## log scale in y-axis
  labs(x = "Genus", y = "Weight (g)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

spp_count_plot <- ggplot(data = yearly_counts, 
                         mapping = aes(x = year, y = n, color = genus)) +
  geom_line() + 
  labs(x = "Year", y = "Abundance")

grid.arrange(spp_weight_boxplot, spp_count_plot, ncol = 2, widths = c(4, 6))
```

```{r}
# grid.arrange() plots
combo_plot <- grid.arrange(spp_weight_boxplot, spp_count_plot, ncol = 2, widths = c(4, 6))

ggsave("combo_plot_abun_weight.png", combo_plot, width = 10, dpi = 300)
```
