---
title: "Learning R with Group - Week 4"
author: "Binod"
date: "4/26/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Covid19 Cases - National

```{r}
library(googlesheets4)
sheets_deauth()

# 
national_covid19 <- read_sheet("1ybnDlI4SiD607WoFTV71OlEsBkLlvzRmhe4Uwam9iNs",
            sheet = "national")

abroad_covid19 <- read_sheet("1ybnDlI4SiD607WoFTV71OlEsBkLlvzRmhe4Uwam9iNs",
            sheet = "abroad")

```
### Data Transformation

```{r}
library(tidyr)

cases_fun <- function(data, cases) {
  data <- data %>%
      pivot_longer(cols = 2:ncol(data),
               names_to = "type" ,
               values_to = "number"
               )
}


national_cases <- cases_fun(national_covid19)

abroad_cases <- cases_fun(abroad_covid19)

```

## Visualize

```{r}
library(ggplot2)
library(dplyr)

con_cases <- national_cases %>% filter(type == "Confirmed")

ggplot(con_cases, aes(x = District, y = number)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = "National Confirmed Cases ", caption = "Data Source: Wikipedia") 
```

```{r}
abroad_con_cases <- abroad_cases %>% filter(type == "Cases")

names(abroad_con_cases) <- c("country", "type", "number")


ggplot(abroad_con_cases, aes(x = country, y = number)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = "Abroad Confirmed Cases ", caption = "Data Source: Wikipedia") 
```

### SAARC 

```{r}

```


## Animal Survey Data set


```{r}
download.file(url="https://ndownloader.figshare.com/files/2292169",
              destfile = "data/survey.csv")
```


Column |	Description
--- | ---
record_id	 | Unique id for the observation
month	| month of observation
day	| day of observation
year	| year of observation
plot_id	| ID of a particular plot
species_id	| 2-letter code
sex	| sex of animal (“M”, “F”)
hindfoot_length	| length of the hindfoot in mm
weight	| weight of the animal in grams
genus	| genus of animal
species	| species of animal
taxon	| e.g. Rodent, Reptile, Bird, Rabbit
plot_type	| type of plot

```{r}
library(readr)

surveys <- read_csv("data/survey.csv")
```

## Structure

```{r}
str(surveys)
```

## Head / TAIL

```{r}
head(surveys)
```

## Summary

```{r}
summary(surveys)
```

## Factorize

```{r}
surveys$sex <- factor(surveys$sex)

?factor
class(surveys$sex)
```


```{r}
surveys
```


### Data Manipulation]

- filter()

# year

```{r}
surveys_1982 <- surveys %>% filter(year == 1982)

species_count <- surveys_1982 %>%
  group_by(species) %>% 
  summarise(count = n())

species_count 

ggplot(species_count , aes(x = species, y = count)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = "Species ", caption = "")

surveys$genus

```

